On the Limitations of Weighted Model Counting

### 2020-03-29 paper.tex
Two more lemmas

\subsection{Lots of Lemmas}

% TODO: maybe 'unite' this with the result about a \/ p, i.e., a ~ a \/ p ~ a /\
% ~p.
% TODO: maybe I should state that A,B are assumed to be BAs with no additional
% properties unless stated otherwise.
\begin{lemma} \label{lemma:elements_in_quotients}
  Let $(p)$ be a principal ideal. Then for any $a \in \mathbf{B}$, $(a \land
  \neg p)/(p) = a/(p)$.
\end{lemma}
\begin{proof}
  Note that
  \[
    (a \land \neg p)/(p) = a/(p) \quad \iff \quad (a \land \neg p) + a \in (p)
    \quad \iff \quad (a \land \neg p) + a \le p.
  \]
  We also have that
  \[
    (a \land \neg p) + a = (a \land \neg p \land \neg a) \lor (\neg(a \land \neg
    p) \land a) = (\neg a \lor p) \land a = (\neg a \land a) \lor (p \land a) =
    p \land a.
  \]
  And, since $p \land a \le p$, we have that $(a \land \neg p) + a \le p$ as
  required.
\end{proof} % TODO: why? (the last bit) It's elementary

\begin{lemma}
  Let $(p)$ be a principal ideal. For any $a, b \in \mathbf{B}$, $a/(p) \le
  b/(p)$ if and only if $a \land \neg p \le b \land \neg p$.
\end{lemma}
\begin{proof}
  Let us begin with the `only of' direction. \Cref{lemma:elements_in_quotients}
  tells us that $(a \land \neg p)/(p) = a/(p)$. Combining this with
  \cref{lemma:canonical_homomorphism,lemma:homomorphisms_and_order} shows that
  \[
    a \land \neg p \le b \land \neg p \quad \implies \quad (a \land \neg p)/(p)
    \le (b \land \neg p)/(p) \quad \iff \quad a/(p) \le b/(p)
  \]
  as required.

@@

\begin{lemma}
  Let $m\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$ be a measure, let $p \in
  \mathbf{B}$ be such that $m(p) = 0$, and let $m^*\colon \mathbf{B}/(p) \to
  \mathbb{R}_{\ge 0}$ be a measure defined by $m^*(a/(p)) = m(a \lor p)$. Then
  for any $a \in \mathbf{B}$, if $a/(p)$ is an atom in $\mathbf{B}/(p)$, then $a
  \land \neg p$ is an atom in $\mathbf{B}$ such that $m^*(a/(p)) = m(a \land
  \neg p)$.
\end{lemma}
\begin{proof} % TODO: This proof is a mess and needs more citations.
  First, we want to show that if $a/(p)$ is an atom, then $a \land \neg p$ is an
  atom. We can instead prove the contrapositive statement, i.e., if there exists
  a $b \in \mathbf{B}$ such that $0 < b < a \land \neg p$, then there exists a
  $b' \in \mathbf{B}$ such that $0/(p) < b'/(p) < a/(p)$. We will show that, in
  fact, we set $b' = b$.
  \Cref{lemma:canonical_homomorphism,lemma:homomorphisms_and_order} already tell
  us that $b/(p) \le a/(p)$, so we only need to show that $0/(p) < b/(p) \ne
  a/(p)$. For the first part, note that
  \[
    0/(p) < b/(p) \quad \iff \quad b/(p) \ne 0/(p) \quad \iff \quad b \not\in
    (p) \quad \iff \quad b \not\le p \quad \iff \quad b \land \neg p \ne 0
  \]
  by \cref{lemma:order}. But if $b \land \neg p = 0$, then $b \land a \land \neg
  p = 0$. This contradicts either that $b \le a \land \neg p$ (i.e., $b \land a
  \land \neg p = b$) or that $b \ne 0$. For the second part, i.e., $b/(p) \ne
  a/(p)$, we will show that if $b/(p) = a/(p)$, and $b \le a \land \neg p$, then
  $b = a \land \neg p$. Indeed,
  \[
    b/(p) = a/(p) \quad \iff \quad a + b \in (p) \quad \iff \quad a+b \le p
    \quad \iff \quad (a+b) \land \neg p = 0,
  \]
  and
  \[
    (a+b) \land \neg p = [(a \land \neg b) \lor (\neg a \land b)] \land \neg p =
    (a \land \neg b \land \neg p) \lor (\neg a \land b \land \neg p),
  \]
  so $(a+b) \land \neg p = 0$ implies that $a \land \neg b \land \neg p = 0$
  which is equivalent to $a \land \neg p \le b$. Therefore we have that $a \land
  \neg p \le b \le a \land \neg p$, so $b = a \land \neg p$ which, by
  contradiction, shows that $b/(p) \ne a/(p)$ and finishes the proof that $0/(p)
  < b/(p) < a/(p)$.

@@

>>>>

\subsection{Lemma Galore}

% TODO: maybe 'unite' this with the result about a \/ p, i.e., a ~ a \/ p ~ a /\
% ~p.
% TODO: maybe I should state that A,B are assumed to be BAs with no additional
% properties unless stated otherwise.
\begin{lemma} \label{lemma:elements_in_quotients}
  Let $(p)$ be a principal ideal. Then for any $a \in \mathbf{B}$, $(a \land
  \neg p)/(p) = a/(p)$.
\end{lemma}
\begin{proof}
  Note that
  \[
    (a \land \neg p)/(p) = a/(p) \quad \iff \quad (a \land \neg p) + a \in (p)
    \quad \iff \quad (a \land \neg p) + a \le p.
  \]
  We also have that
  \[
    (a \land \neg p) + a = (a \land \neg p \land \neg a) \lor (\neg(a \land \neg
    p) \land a) = (\neg a \lor p) \land a = (\neg a \land a) \lor (p \land a) =
    p \land a.
  \]
  And, since $p \land a \le p$, we have that $(a \land \neg p) + a \le p$ as
  required.
\end{proof} % TODO: why? (the last bit) It's elementary
% TODO: remove all those 'let (p) be a principal ideal'.

\begin{lemma} \label{lemma:min_element_and_order}
  Let $(p)$ be a principal ideal. For any $a, b \in \mathbf{B}$, $a/(p) \le
  b/(p)$ if and only if $a \land \neg p \le b \land \neg p$.
\end{lemma}
\begin{proof}
  Let us begin with the `only of' direction. \Cref{lemma:elements_in_quotients}
  tells us that $(a \land \neg p)/(p) = a/(p)$. Combining this with
  \cref{lemma:canonical_homomorphism,lemma:homomorphisms_and_order} shows that
  \[
    a \land \neg p \le b \land \neg p \quad \implies \quad (a \land \neg p)/(p)
    \le (b \land \neg p)/(p) \quad \iff \quad a/(p) \le b/(p)
  \]
  as required.

@@

\begin{lemma} \label{lemma:measure_of_atom}
  Let $m\colon \mathbf{B} \to \mathbb{R}_{\ge 0}$ be a measure, let $p \in
  \mathbf{B}$ be such that $m(p) = 0$, and let $m^*\colon \mathbf{B}/(p) \to
  \mathbb{R}_{\ge 0}$ be a measure defined by $m^*(a/(p)) = m(a \lor p)$. Then
  for any $a \in \mathbf{B}$, if $a/(p)$ is an atom in $\mathbf{B}/(p)$, then $a
  \land \neg p$ is an atom in $\mathbf{B}$ such that $m^*(a/(p)) = m(a \land
  \neg p)$.
\end{lemma}
\begin{proof} % TODO: This proof is a mess and needs more citations.
  First, we want to show that if $a/(p)$ is an atom, then $a \land \neg p$ is an
  atom. We can instead prove the contrapositive statement, i.e., if there exists
  a $b \in \mathbf{B}$ such that $0 < b < a \land \neg p$, then there exists a
  $b' \in \mathbf{B}$ such that $0/(p) < b'/(p) < a/(p)$. We will show that, in
  fact, we set $b' = b$.
  \Cref{lemma:canonical_homomorphism,lemma:homomorphisms_and_order} already tell
  us that $b/(p) \le a/(p)$, so we only need to show that $0/(p) < b/(p) \ne
  a/(p)$. For the first part, note that
  \[
    0/(p) < b/(p) \quad \iff \quad b/(p) \ne 0/(p) \quad \iff \quad b \not\in
    (p) \quad \iff \quad b \not\le p \quad \iff \quad b \land \neg p \ne 0
  \]
  by \cref{lemma:order}. But if $b \land \neg p = 0$, then $b \land a \land \neg
  p = 0$. This contradicts either that $b \le a \land \neg p$ (i.e., $b \land a
  \land \neg p = b$) or that $b \ne 0$. For the second part, i.e., $b/(p) \ne
  a/(p)$, we will show that if $b/(p) = a/(p)$, and $b \le a \land \neg p$, then
  $b = a \land \neg p$. Indeed,
  \[
    b/(p) = a/(p) \quad \iff \quad a + b \in (p) \quad \iff \quad a+b \le p
    \quad \iff \quad (a+b) \land \neg p = 0,
  \]
  and
  \[
    (a+b) \land \neg p = [(a \land \neg b) \lor (\neg a \land b)] \land \neg p =
    (a \land \neg b \land \neg p) \lor (\neg a \land b \land \neg p),
  \]
  so $(a+b) \land \neg p = 0$ implies that $a \land \neg b \land \neg p = 0$
  which is equivalent to $a \land \neg p \le b$. Therefore we have that $a \land
  \neg p \le b \le a \land \neg p$, so $b = a \land \neg p$ which, by
  contradiction, shows that $b/(p) \ne a/(p)$ and finishes the proof that $0/(p)
  < b/(p) < a/(p)$.

@@

\begin{lemma} \label{lemma:minimal_in_quotient} % TODO: define completeness
  Let $\mathbf{B}$ be a complete BA. For any $a, b \in \mathbf{B}$, if $a/(p) =
  b/(p)$, then $a \land \neg p = b \land \neg p$. As a consequence, $a \land
  \neg p \le b$.
\end{lemma}
\begin{proof} % TODO: rename b to something else
  As in the proof of \cref{lemma:measure_of_atom}, $a/(p) = b/(p)$ implies that
  $a \land \neg p \le b$. Since $\mathbf{B}$ is complete, let $b = \bigwedge \{
  c \in \mathbf{B} \mid c/(p) = a/(p) \}$; then we still have that $b/(p) =
  a/(p)$. But then $b \le a \land \neg p \le b$, so $a \land \neg p = b$. This
  defines $a \land \neg p$ independently of $a$ as the least element in $\{c \in
  \mathbf{B} \mid c/(p) = a/(p) \}$.
\end{proof}

\begin{corollary}
  For any complete BA $\mathbf{B}$, if $a \in \mathbf{B}$ is an atom, then
  $a/(p)$ is either an atom or $0/(p)$. In the former case, $a = a \land \neg
  p$.
\end{corollary}
\begin{proof}
  Since \cref{lemma:minimal_in_quotient} tells us that for all $b \in
  \mathbf{B}$, if $b/(p) = a/(p)$, then $b \ge a \land \neg p$, if there is an
  atom $b \in \mathbf{B}$ such that $b/(p) = a/(p)$, then it must be $a \land
  \neg p$. If $a$ is an atom, then $a \land \neg p \le a$ implies that either $a
  = a \land \neg p$ or $a \land \neg p = 0$. The latter is equivalent to $a/(p)
  = 0/(p)$ by \cref{lemma:order}. The former, combined with the assumption that
  $a$ is an atom and \cref{lemma:min_element_and_order}, implies that $a/(p)$ is
  an atom.
\end{proof}
